---
title: "Determine Error"
author: "Bowen Jones"
date: "9/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dglm)
```

In `dglm()` call for df.breed data, error returned when fitting dispersion model.
```{r}
df.breed <- read.csv(file = "simdata3.csv")
df.breed <- df.breed[-c(1,2),]

sample <- c(round(runif(10, 1,900),0))

mod = dglm(stress ~ ., dformula = ~., data = df.breed, subset = sample)
```

After reading the error output, I speculate that it is due to repeated stress values.
```{r}
length(unique(df.breed$stress))==length(df.breed$stress)
```

What if we add jitter to create fully unique values?
```{r}
df.breed$stress <- jitter(df.breed$stress)
length(unique(df.breed$stress))==length(df.breed$stress)

mod = dglm(stress ~ ., dformula = ~., data = df.breed[sample,])
```

Let's increase jitter.
```{r}
df.breed$stress <- jitter(df.breed$stress, amount = 1)
mod = dglm(stress ~ ., dformula = ~., data = df.breed[sample,])
```

Again?
```{r}
df.breed$stress <- jitter(df.breed$stress, amount = 5)
mod = dglm(stress ~ ., dformula = ~., data = df.breed[sample,])
```



Hmm. Let's see if we can reproduce this error. Start by simulating data.
```{r}
dat <- cbind(round(rnorm(100, 38, 2),0), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5), rbinom(100,1,.5))
dat <- as.data.frame(dat)

length(unique(dat$V1))==length(dat$V1)
```

```{r}
sample2 <- c(round(runif(10, 1,100),0))

m <- dglm(V1 ~ ., dformula = ~.,  data = dat[sample2,])
```
Cool, got the error.

Jitter?
```{r}
dat$V1 <- jitter(dat$V1, amount = 1)
length(unique(dat$V1))
m <- dglm(V1 ~ ., dformula = ~.,  data = dat[sample2,])
```

What if fewer predictors?
```{r}
data <- dat[,1:5]

m <- dglm(V1 ~ ., dformula = ~.,  data = data[sample2,])
```
Great! Can we just change the formula?
```{r}
m <- dglm(V1 ~ ., dformula = ~ V2+V3+V4+V5, data = dat[sample2,])
```


Differences in formula?
```{r}
m <- dglm(V1 ~ V2+V3+V4+V5, dformula = ~ V2+V3+V4+V5, data = dat[sample2,])
```
